---
title: "STAT 230 HW 1"
author: "Owen Forman" 
output:
  pdf_document: default
  html_document:
    df_print: paged
---

**1.** (Conceptual review) Chapter 2, exercises 3, 4, and 7

The answers to the following questions can be found on page 56. You should think about these questions before looking at the answers. You will receive full credit for answering the questions.

\textbf{Chapter 2, Exercise 3 Answer:}

\vspace{4pt}

\textbf{Chapter 2, Exercise 4 Answer:}

\vspace{4pt}

\textbf{Chapter 2, Exercise 7 Answer:}

\vspace{15pt}

**2.** (Grizzlies)

Load the following dataset:

```{r}
grizzly <- read.csv('https://www.math.carleton.edu/ckelling/data/Grizzly.csv', header=TRUE)
```

The data record the length and girth measurements (in inches) for a sample of grizzlies.

(a) Plot the bear's girth (response variable) against its length (predictor variable) and show the least squares line. Make sure the plot has appropriate axis labels. Comment on the apparent relationship (strength, shape, direction).

\textbf{Answer:} The relationship appears to have a relatively strong positive correlation. The relationship also has

```{r}
#code with comments here

#load libraries
library(ggplot2)

#creates base graph (girth against length) with points shown
bear_girth_on_length <- ggplot(data=grizzly) +aes(x = Length, y = Girth) + geom_point() + labs(title = "regression of bear girth against length", x = "Length of Bear (in inches)", y = "Girth of Bear (in inches)")

#adds least squares line
bear_girth_on_length + geom_smooth(method = "lm", se = FALSE)
```

(b) Find (and write out) the equation of the estimated linear model of the expected girth given length. (Here I am just asking for the equation of the fitted values, $\hat{Y}$; it is not necessary to state the assumptions about the errors here. Do not just print the R output; you need to write down the equation.)

\textbf{Answer:} $\hat{Y}$ = 32.6678 + .5462(Length)

(Note that in the equation $\hat{Y}$ is the predicted Girth of the grizzly)

```{r}
#code with comments here
fit <- lm(Girth ~ Length, data = grizzly)
fit

```

(c) What is the interpretation of the slope coefficient?

\textbf{Answer:} "for every extra inch of length, we expect the girth of the bear to increase by .5462 inches"

(d) Give and interpret a 95% confidence interval for the slope coefficient. (You can use the `confint()` command.)

\textbf{Answer:} We are 95% confident that for every inch of length longer a Bear is, the girth of the bear increases by 

```{r}
#code with comments here

confint(fit)
```

(e) Suppose for a different species of bear, the scientists sample 25 bears and for their estimated linear model they find a slope coefficient estimate of $\hat{\beta} = 0.425$ and SE($\hat{\beta}$) = 0.068. Give a 90% confidence interval for the true slope coefficient. (Note: this question does not rely on the dataset used for the questions above. Use the estimated slope and SE given here, and the `qt()` function to get the appropriate critical value.)

\textbf{Answer:}

```{r}
#code with comments here

```

## Recommended conceptual exercises (do not turn in, answers in book/Moodle):

Chapter 7 exercises 3, 4, 5, 9, 10, 12
